---
title: "Kaneohe Bay Reef Summary"
author: "Nan Himmelsbach"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data import, echo=FALSE, results='hide', include=FALSE}
################################################
# import data, observe structure
################################################
data<-read.csv("~/Desktop/kbayreef2/kbayreef/groupdata.csv", header = TRUE)
names(data)
head(data)
data$year<-substr(data$date,0,4) # add a column for year only
```

## Introduction

This study aims to explore various methods of measuring patch reefs in Kaneohe Bay, Hawai'i. Through the use of satellite imagery via GoogleEarth and Planet Explorer

## Methods

**Study Site**
Reefs were selected within Kaneohe Bay, Hawai'i for measurement. All reefs were circular and reefs were classifed by size: small (patch reef 19), medium (patch reef 16), large (patch reef 12).

**Field Reef Measurements**
To measure the perimeter, reefs were encircled via snorkeling. The tape was laid on the reef just above the point where the gradient of the reef's slope was steepest. 

To measure the area, snorkelers swam over the reef once in a north-south direction, followed by an east-west direction to measure the maximum and minimum diameter of the reef. 

**Satellite Imagery Measurements**
Reefs were measured for perimeter and area using Google Earth and Planet Explorer. In Google Earth, the resolution of imagery data used included 0.15, 0.5, 1.0, and 2.0 m. 


## Results

```{r data explore-1, echo=FALSE, results='hide', fig.height=5, fig.width=5, fig.align='center'}
################################################
# explore patterns
################################################
# Effect of method
plot(area_m2~method, data=data, na.ignore=TRUE, col="lightgrey",
     ylab="Patch reef area (m^2)", 
     xlab="Method", 
     main="Area versus method")
dev.copy(pdf,"figs/method_area_effect.pdf")    # can change to .png, etc.; can change size to incr resolution
dev.off()

# Effect of imagery resolution
plot(area_m2~imagery_resolution_m, data=data, col="slategrey",
     xlim=c(0,10),
     xlab="Imagery spatial resolution (m)", 
     ylab="Patch reef area (m^2)", 
     main="Area versus imagery resolution")
res.area.lm=lm(area_m2 ~ imagery_resolution_m, data=data)
abline(res.area.lm, col = "black")
summary(res.area.lm)
dev.copy(pdf,"figs/resolution_area_effect.pdf")    
dev.off()
plot(perimeter_m~imagery_resolution_m, data=data, col="slategrey",
     xlim=c(0,10),
     xlab="Imagery spatial resolution (m)", 
     ylab="Patch reef perimeter (m)", 
     main="Perimeter versus imagery resolution")
res.perim.lm=lm(perimeter_m ~ imagery_resolution_m, data=data)
abline(res.perim.lm, col = "black")
summary(res.perim.lm)
dev.copy(pdf,"figs/resolution_perimeter_effect.pdf")    
dev.off()
# Effect of date (i.e., do patch reefs measureably grow or shrink over time?)
plot(area_m2~as.numeric(year), data=data, col="slategrey",
     xlab="Year", 
     ylab="Patch reef area (m^2)", 
     main="Area versus year")
date.area.lm=lm(area_m2 ~ as.numeric(year), data=data)
abline(date.area.lm, col = "black")
summary(date.area.lm)
dev.copy(pdf,"figs/date_area_effect.pdf")    
dev.off()
```

The data suggest that the field measurements are more accurate than the satellite imagery measurements (Google Earth and Planet Explorer combined).


Now, let's see if there's anything going on with time when imagery resolution is considered:

```{r data analysis-1, echo=FALSE, results='hide', fig.height=5, fig.width=5, fig.align='center'}
################################################
# analyze data
################################################
# Effects of imagery resolution + year
res.year.lm=lm(area_m2 ~ imagery_resolution_m * as.numeric(year), data=data)
summary(res.year.lm)
```

## Discussion


